---
title: "tiledb-intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tiledb-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
# library(tiledb)
library(data.table)
library(lubridate)
```



```{r}
# # creditentials for AWS S3
# config <- tiledb_config()
# config["vfs.s3.aws_access_key_id"] <- Sys.getenv("AWS-ACCESS-KEY")
# config["vfs.s3.aws_secret_access_key"] <- Sys.getenv("AWS-SECRET-KEY")
# config["vfs.s3.region"] <- Sys.getenv("AWS-REGION")
# context_with_config <- tiledb_ctx(config)
```



```{r}
# # import aapl stock
# arr <- tiledb_array("s3://equity-usa-minute-fmpcloud", 
#                     as.data.frame = TRUE)
# system.time(spy <- arr["SPY"])
# 
# # inspect
# head(spy)
```

## I like to work with dataframes (optional)

```{r}
# spy <- as.data.table(spy)
```


## Timezones in UTC

```{r}
# spy[, time := as.POSIXct(time, origin = as.POSIXct("1970-01-01 00:00;00", tz = "UTC"))]
```

## Note on time zones

We can check timezone by looking at atributes of an object.

```{r}
# # define POSIXct without timezone
# x <- as.POSIXct("2022-01-01 15:30:20") 
# attributes(x)
# # LOCAL TIMEZONE
# 
# # define POSIXct without timezone
# with_tz(x, tzone = "UTC")
# with_tz(x, tzone = "CET")
# with_tz(x, tzone = "Europe/Zagreb")
# 
# # Recomendation
# # 1. Always define timezone
# x <- as.POSIXct("2022-01-01 15:30:20", tz = "UTC") 
# with_tz(x, tzone = "UTC")
# with_tz(x, tzone = "CET")
# with_tz(x, tzone = "Europe/Zagreb")
# # 2. If you plan to work with UTC most times define environemnt variable TZ
# Sys.setenv(TZ="UTC") # or define in .Renviron file
# 
# 
# # change timezone
# # 1. use lubirdate package
# x <- as.POSIXct("2022-01-01 15:30:20", tz = "UTC") 
# print(x)
# with_tz(x, tzone = "America/New_York")
```

One important note. If ypou misspell timezone, it will convert it to GMT! Beforelatest version of R this was done without warning which was inconvinient. Now we get an warning. Example:

```{r}
# x <- as.POSIXct("2022-01-01 15:30:20", tz = "America/New_Yrk") 
# print(x)
```


When you change timezone, you change only the representation of the time zone (format), but not the actual numeric that represents the timezone, which is unchanged. Here is an example:

```{r}
# x <- as.POSIXct("2022-01-01 15:30:20", tz = "UTC")


```


## FirstRate data

```{r}
# arr <- tiledb_array("s3://equity-usa-minute-firstrate-adjusted", 
#                     as.data.frame = TRUE)
# system.time(aapl <- arr["AAPL"])
# 
# # inspect
# head(aapl)
```

## Timezone

```{r}

```



